---
title: "Data Manipulation"
output: github_document
---

This file is for doing data manipulation. 

```{r}
library(tidyverse)
```

```{r}
litters_df = 
  read_csv(
    "data/FAS_litters.csv",
    na = c("", "NA", ".")
  )

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv(
    "data/FAS_pups.csv",
    skip = 3, 
    na = c("", "NA", ".")
  )

pups_df = janitor::clean_names(pups_df)
```

### `select` some variables

```{r}
select(litters_df, group:gd_of_birth)

select(litters_df, starts_with("gd"), group)

select(litters_df, GROUP = group, everything())

select(litters_df, -starts_with("gd"))

relocate(litters_df, litter_number)
```

In the pups data, select litter number, sex, and PD ears. 

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` some rows

```{r}
filter(litters_df, gd_of_birth == 20)

filter(litters_df, gd0_weight != 0)

filter(litters_df, group %in% c("Con7", "Con8"))

drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows PD walk is less than 11 and sex is 2. 

```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```

## `mutate` to change some columns
```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain ^ 2,
       group = str_to_lower(group)
       )
```

In the pups datast, create a variable which subtracts 7 from pd walk. 

```{r}
mutate(pups_df, 
       pd_walk_minus_7 = pd_walk - 7
       )
```

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(gd0_weight))
```

## Do multiple steps 
```{r}
litters_df = read_csv(
  file = "data/FAS_litters.csv",
  na = c("", "NA", ".")
  ) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight, 
    group = str_to_lower(group)
  )
```

Load pups, clean names, drop missing, keep litter number and pd variables, add pd walk minus 7

```{r}
pups_df = 
  read_csv(
    "data/FAS_pups.csv",
    skip = 3, 
    na = c("", "NA", ".")
  ) |> 
  janitor::clean_names() |> 
  drop_na() |> 
  select(-sex) |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
    )
```

Sometimes the first thing is not a dataframe. 
```{r}
litters_df |> 
  lm(gd18_weight ~ gd0_weight, data = _)
```

```{r}
litters_df$group
```



